TechHacks Talk - Dec. 6, 2013 - Recommender System
========================================================
rec-sys : : [HTML](http://www.cs.mtu.edu/~lebrown/TechHacks-f13/rec-sys/rec-sys.html)  | [Rmd - R Markdown](http://www.cs.mtu.edu/~lebrown/TechHacks-f13/rec-sys/rec-sys.Rmd) | [md - Markdown](http://www.cs.mtu.edu/~lebrown/TechHacks-f13/rec-sys/rec-sys.md)


## Data Mining Example - Recommender Systems


We will explore building a recommender system.  

### Load data

First step is to load in some sample data.  The data was collected by survey from cs1000 students in Fall 2013 and is available [here](http://www.cs.mtu.edu/~lebrown/TechHacks-12-6-13/rec-sys/data/cs1000_f13_data.csv)  

```{r}
dat1 <- read.csv("~/Desktop/TechHacks-12-6-13/data/cs1000_f13_data.csv")
```
Note, you will need to set the data to load from your working directory or where it is located on your computer. 


### Install and Load RecommenderLab R package

We will make use of some available R functions.

If you haven't already installed the Recommender Lab package, do so:
```{r, eval=FALSE}
install.packages('recommenderlab', dependencies=TRUE)
```

Then load the package:
```{r, include=FALSE}
library(recommenderlab)
```

### Work with Data

First, set up the data in the needed format for the recommender library functions. 

```{r}
m1 <- as.matrix(dat1)
r1 <- as(m1, "realRatingMatrix")
```

In order, to get a view on the distribution of ratings we can plot the ratings as heat maps.

```{r ratingPlot, fig.width=6}
rn1 <- normalize(r1)
image(r1, main="Ratings");
image(rn1, main="Normalized Ratings");
```

Next, we can look at the mean rating for each student and plot this as a histogram.

```{r}
rmn <- rowMeans(r1)
rmn
hist(rmn, breaks=10, main="Histogram of Mean User Ratings", xlab="Ratings")
```

We can also look at the mean rating for each show.
```{r}
cmn <- colMeans(r1)
cmn
hist(cmn, breaks=10, main="Histogram of Mean TV Rating", xlab="Ratings")
```

In fact, we can see by sorting via the mean TV rating, that the highest rated shows are: 

* Breaking Bad
* The Walking Dead
* Game of Thrones

```{r}
nd <- order(cmn)
colnames(dat1)[nd]
```

----
Following examples of recommenderlab documentation.
### Recommendation System

```{r}
ubr <- Recommender(r1, method="UBCF")
pred <- predict(ubr, r1, type="ratings")
#as(pred, "matrix")
```

The supplied ratings for User 3 are:
```{r}
dat1[3,]
```

The predicted ratings for the missing shows are:
```{r}
colnames(dat1)[is.na(dat1[3,])]
getRatings(pred[3,])
```


Present the top 2 predicted ratings for each user (output surpressed for space)
```{r, eval=FALSE}
pred <- predict(ubr, r1, n=2)

recs <- bestN(pred, n=2)
as(recs, "list")
```

Present the top 3 predicted ratins for user 3
```{r}
pred <- predict(ubr, r1[3], n=3)
as(pred, "list")
```